---
title: "Assignment 1"
author: "Jinjie Zhang"
date: "October 4, 2017"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Section 4.1.1
### Exercise 2
(a)
```{r chunk1}
Fibonacci=numeric(31)
Fibonacci[1]=Fibonacci[2]=1
for(i in 3:31){
  Fibonacci[i]=Fibonacci[i-1]+Fibonacci[i-2]
}

 ratio=numeric(30)              #construct the sequence of ratios
 for (i in 1:30){
   ratio[i]=Fibonacci[i+1]/Fibonacci[i]
 }
 ratio
 
```
When $n$ is large, the value of $f_n/f_{n-1}$ is close to $1.618034$. It seems that the sequence $\{f_n/f_{n-1}\}$ is convergent.

(b)
```{r chunk2}
G_ratio=(1+sqrt(5))/2
G_ratio
abs(G_ratio-ratio[30])
```
We find that the $f_n/f_{n-1}$ is pretty close to $\frac{1+\sqrt{5}}{2}$ as $n$ goes to infinity. In fact, we can prove this result rigorously, see [this](http://mathworld.wolfram.com/FibonacciNumber.html).

### Exercise 3
(a)
The answer should be $\sum_{j=1}^{5}j=15$. 
```{r chunk3,echo=TRUE}
answer <- 0
for (j in 1:5) answer <- answer + j
answer
```
(b)
The answer should be the collection $(1,2,3,4,5)$.
```{r chunk4,echo=TRUE}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
answer
```

(c)
The answer should be the collection $(0,1,2,3,4,5)$
```{r chunk5,echo=TRUE}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
answer
```
(d)
The answer should be the factorial $5!=120$.
```{r chunk6,echo=TRUE}
answer <- 1
for (j in 1:5) answer <- answer * j
answer
```
(e)
```{r chunk7,echo=TRUE}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
answer
```
Yes. We can predict successive numbers accroding to its periodicity. The following elements should be $21,23,6,11\dots$

## Section 4.1.2
### Exercise 4
```{r chunk8,echo=TRUE}
interest=function(P,n){
  #P is the initial investment amount; n is the number of interest periods(years);
  #I is the amount of interest earned  
  if (n%%1!=0|n<0){
    print('the year should be a positive integer!')
  }
  else if(n<=3){
    I=P*((1+0.04)^n-1)
    return(I)
  }
  else if(n>3){
    I=P*((1+0.05)^n-1)
    return(I)
  }
}
```
### Exercise 5
```{r chunk9,echo=TRUE}
mortgage=function(P,n,open){
  if (n%%1!=0|n<0){
    print('the year shoud be a positive integer!')
  }
  else if(open==TRUE){
    i=0.005
    R=P*i/(1-(1+i)^(-n))
    return(R)
  }
  else if(open==FALSE){
    i=0.004
    R=P*i/(1-(1+i)^(-n))
    return(R)
  }
}
```

## Section 4.1.3
### Exercise 2
```{r chunk10,echo=TRUE}
Fibonacci=c(1,1)
while(sum(tail(Fibonacci,2)) < 300){
  Fibonacci=c(Fibonacci,sum(tail(Fibonacci,n=2)))
}
Fibonacci
```

### Exercise 4
```{r chunk11}
error=1e-6
value_old=0.006
value_new=0
count=0
while(abs(value_new-value_old)>error){
  if(count>0){
    value_old=value_new
  }
  value_new=(1-(1+value_old)^(-20))/19
  count=count+1
}
value_new
count  #the number of iterations
```
If the initial value is $0$, then the output is $0$. If we try other proper initial value (around $0.01$), the results may have small difference. 

## Section 4.1.5
###Exercise 2
(a) This algorithm finds all prime number from $2$ to $n$. In each step, we remove the numbers that is divided by the smallest remaining number in sequence 'sieve'. We classify the smallest remaining number as a prime and continue this process until all numbers in 'sieve' are removed.
(b) Suppose that integer $m\in[2,n]$ has decompostion $n=pq$ with $p>\sqrt{n}$. Then we must have $q<\sqrt{n}$ and thus $m$ should be removed before $p$ increases to current value. According to this kind of symmetry, we only need to check the case where $p\leq\sqrt{n}$.  

(c)
```{r chunk12}
Eratosthenes <- function(n) {
 # Print prime numbers up to n (based on the sieve of Eratosthenes)
 if (n >= 2) {
 sieve <- seq(2, n)
 primes=c()
 while (length(sieve) > 0) {
 p <- sieve[1]
 primes <- c(primes, p)
 sieve <- sieve[(sieve %% p) != 0]
 if(p>sqrt(n)){
   break
 }
 }
 primes <- c(primes, sieve)
 return(primes)
} 
  else {
  stop("Input value of n should be at least 2.")
 }
}

#Example 
Eratosthenes(30)
```

## Section 4.2.1
###Exercise 2
(a)
```{r}
compound.interest=function(P,i.r,n){
  Total=P*(1+i.r)^n
  return(Total)
}

```
(b)
```{r}
compound.interest(1000,0.01,30)

```
Mr.Ng will have $1347.849 in the bank at the end of 30 months.

###Exercise 3
```{r}
#Find root of f(x)=0 using bisection method
Bisection=function (f, a, b, num, eps = 1e-05) 
{
    h = abs(b - a)/num
    i = 0
    j = 0
    a1 = b1 = 0
    while (i <= num) {
        a1 = a + i * h
        b1 = a1 + h
        if (f(a1) == 0) {
            print(a1)
            print(f(a1))
        }
        else if (f(b1) == 0) {
            print(b1)
            print(f(b1))
        }
        else if (f(a1) * f(b1) < 0) {
            repeat {
                if (abs(b1 - a1) < eps) 
                  break
                x <- (a1 + b1)/2
                if (f(a1) * f(x) < 0) 
                  b1 <- x
                else a1 <- x
            }
            j = j + 1
            print((a1 + b1)/2)
        }
        i = i + 1
    }
    if (j == 0) 
        print("finding root is fail")
    else print("finding root is successful")
  }


#Example: let f(x)=x^2-e^x
f=function(x){
  value=x^2-exp(x)
  return(value)
}
Bisection(f,-2,2,20)
uniroot(f,c(-1,1))$root #compare our result with the true value

```
We find that our result is pretty close to the true value and we can plot the graph of this function on interval $[-2,2]$.
```{r}
plot(f,-2,2)
abline(h=0,v=0,col="blue",lty=2)
```

##section 4.4.1
###Exercises 1
The modified code is as follows.
```{r}
mergesort = function (x, decreasing=FALSE) {
n = length(x)
if(n < 2){
seq = x
}
else{
m = floor(n/2)
y = x[1:m]
z = x[(m+1):n]
y = mergesort(y)
z = mergesort(z)
seq = c()
while(min(length(y),length(z)) > 0) {
if(y[1] < z[1]) {
seq = c(seq, y[1])
y = y[-1]
} else {
seq = c(seq, z[1])
z = z[-1]
}
}
if(length(y) > 0)
seq = c(seq, y)
else
seq = c(seq, z)
}
if(decreasing==TRUE)
return(rev(seq))
else
return(seq)
}

#We give an example
x=rnorm(5)
x
mergesort(x,FALSE)   #sort in increasing order
mergesort(x,TRUE) #sort in decreasing order
```
###Exercise 2
(a)
Use Newton's method to solve equations.
```{r}
Newton_method= function(x_initial, y_initial, f, g, tol=1e-5) {
df = deriv(f, c("x","y"))
dg = deriv(g, c("x","y"))  #find the partial derivatives
x = x_initial
y = y_initial
while(abs(eval(f)) > tol | abs(eval(g)) > tol) {
evdf = eval(df)
evdg = eval(dg)
f = evdf[1]
fx = attr(evdf,"gradient")[1]
fy = attr(evdf,"gradient")[2]
g = evdg[1]
gx = attr(evdg,"gradient")[1]
gy = attr(evdg,"gradient")[2]
d = fx * gy - fy * gx
x = x - (gy*f - fy*g)/d  #get new x and new y
y = y - (fx*g - gx*f)/d
}
return(c(x,y))
  
}
```
(b)
```{r}
f = expression(x + y)
g = expression(x^2 + 2*y^2 - 2)
Newton_method(0.5,-0.5,f,g)
Newton_method(-0.5,0.5,f,g)
```
In fact, the analytic solution of this equation system is $(\sqrt{\frac 2 3},-\sqrt{\frac 2 3})$ and $(-\sqrt{\frac 2 3},\sqrt{\frac 2 3})$. Our result is pretty close to the true value.

##Chapter 4
###Exercises 1
```{r}
directpoly=function(x,coef){
  degree=length(coef)
  value=0
  for(i in 1:degree){
    value=value+coef[i]*x^(i-1)
  }
  return(value)
}

```
###Exercise 2
```{r}
hornerpoly = function(x,coef) {
n = length(coef)
a = matrix(0, nrow=length(x), ncol=n)  #construc a matrix 
a[,n] = coef[n]
for(i in (n-1):1) {
a[,i] = a[,i+1] * x + coef[i]
}
return(a[,1])
}
#Give an example
hornerpoly(c(2,4,6),c(10,2,3,4,5))

```
###Exercise 3
(a)
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000),c(1, -2, 2, 3, 4, 6, 7)))
```

```{r}
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(1, -2, 2, 3, 4, 6, 7)))

```
We notice that 'hornerpoly' function computes faster than 'directpoly'.

(b)
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000),c(-3,17,2)))
```

```{r}
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(-3,17,2)))

```
Remark: The difference of time consumption is smaller and 'hornerpoly' function is still faster than 'directpoly'.
